---
title: "Predicted seagrass carbon map at Ria de Aveiro for 2024 and 2025"
author: "Lee, Chengfa Benjamin"
date: "`r Sys.Date()`"
output: bookdown::html_document2
bibliography: references.bib 
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(magrittr)
library(sf)
library(dplyr)
library(leaflet)
library(terra)
# library(readxl)
# library(tidyr)
# library(ggplot2)
# library(ggiraph)
# library(ggpubr)
# library(patchwork)
```

This page documents the results of the _Zostera noltii_ seagrass carbon mapping work at Ria de Aviero. This includes work done in:

* [The fieldtrip of Aveiro 2025](https://sigoiry.github.io/Spectra_Aveiro_2025/) by [Simon Oiry](https://github.com/SigOiry)
* [Coring Carbon Seagrass](https://sigoiry.github.io/Coring_Carbon_Seagrasses/) by Simon Oiry
* [Exploring the relationship between seagrass reflectance and biomass](https://leechengfa.github.io/Seagrass_reflectance_biomass_2025.github.io/)

All analyses were performed in R version 4.5.2 [@Rversion].
All work belong to the [REWRITE project](https://rewriteproject.eu/).

# Study Location

This page focuses on Ria de Aveiro, a [demonstrator site for the REWRITE PROJECT](https://rewriteproject.eu/demonstrator-network/ria-de-aveiro). A natural site and rewilded site was designated in the lagoon and seagrass meadows were found in both sites. Its seagrass meadow was estimated to be 3.52 km^2^ during the peak season of 2024 [@davies2024].

The sites were sampled according to a grid, and additional stations were sampled randomly within the grid.

```{r samplingDesign}
#| echo: false
#| error: false
#| message: false
#| warning: false

stations <- read_sf('data/shp/Points.shp') %>%
  filter(!is.na(seagrassA)) %>%
  select(name, type, site, habitat, longitude, latitude)

map_centre <- stations %>% 
  st_union() %>%
  st_centroid() %>%
  st_coordinates()

pal.shp <- colorFactor(c('darkgreen','pink4'), domain = c('natural','rewilded'))

m.shp <- leaflet(stations) %>%
  addTiles() %>%
  setView(map_centre[1], map_centre[2], 14) %>%
  addCircleMarkers(radius = 2, color = ~pal.shp(site), fillOpacity = 1, label = ~name) %>%
  addLegend(pal = pal.shp, values = ~site, opacity = 1)

m.shp
```

# ICE CREAMS Model prediction

ICE CREAMS classification

```{r icecreams}
#| echo: false
#| error: false
#| message: false
#| warning: false

import.icecreams <- function(image.path) {
  map.ice <- rast(image.path)
    map.ice[map.ice == 0 | map.ice == -1] <- NA
    
  return(map.ice)
}

pal.icecreams <- c('khaki','slategrey','lightgreen','darkgreen','aquamarine','brown','red','darkblue','yellowgreen')
labels.icecreams <- c("Bare Sand","Bare Sediment","Chlorophyta","Magnoliopsida","Microphytobenthos","Phaeophyceae","Rhodophyta","Water","Xanthophyceae")

map.class <- import.icecreams('data/ice_creams_prediction/S2C_MSIL2A_20250715T113341_N0511_R080_T29TNF_20250715T165315_Pred_BFRD_V1_3.tif')
  levels(map.class$Out_Class) <- data.frame(id=1:9, cover = labels.icecreams)

pal.class <- colorFactor(pal.icecreams, domain = 1:9, na.color = "transparent")

m.class <- leaflet() %>%
  addTiles() %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  addRasterImage(map.class, colors = pal.class) %>%
  addLegend(colors = pal.icecreams, values = 1:9, labels = labels.icecreams, opacity = 1)
m.class
```

Sentinel-2 NDVI Map

```{r ndvi}
#| echo: false
#| error: false
#| message: false
#| warning: false

import.ndvi <- function(image.path) {
  map.index <- rast(image.path)
  
  return(map.index)
}

map.ndvi <- import.ndvi('data/ndvi/S2C_MSIL2A_20250715T113341_N0511_R080_T29TNF_20250715T165315_ndvi.tif')

pal.ndvi <- colorNumeric(c('red','white','green'), domain = c(-1, 0, 1), na.color = "transparent")

m.ndvi <- leaflet() %>%
  addTiles() %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  addRasterImage(map.ndvi, colors = pal.ndvi) %>%
  addLegend(pal = pal.ndvi, values = c(-1, 0, 1), opacity = 1)
m.ndvi
```

# Biomass prediction 

Rewilded vs Natural, Averaged results for the mapping

```{r predictedbiomass}
#| echo: false
#| error: false
#| message: false
#| warning: false

derive.agb <- function(ndvi) {
  agb <- ndvi * 140 - 9.5
    agb[agb < 0] <- 0
  
  return(mask(agb, map.class$Out_Class, inverse = TRUE, maskvalues = 4))
}

map.agb <- derive.agb(map.ndvi)

derive.bgb <- function(agb) {
  bgb <- agb * 0.69 + 9.1
    bgb[agb == 0] <- 0
  
  return(bgb)
}

map.bgb <- derive.bgb(map.agb)

## calculate AGB and BGB over area --> sum? // take care of units (g/m2) // how to account for this with S2?? (10m not 1m // 100m2 not 1m2)
## conversion to tons/ha (multiply by 10 and divide by 1000)
biomass.agb <- global(map.agb, fun = 'sum', na.rm = TRUE) * 10 / 1000
biomass.bgb <- global(map.bgb, fun = 'sum', na.rm = TRUE) * 10 / 1000

## shared palette
pal.biomass <- colorNumeric(c('white','black'), domain = c(0, 70), na.color = 'transparent')
```

Total AGB in kg/ha: `r round(biomass.agb$sum, 2) %>% format(scientific = FALSE)`

```{r mapAGB}
#| echo: false
#| error: false
#| message: false
#| warning: false

m.agb <- leaflet() %>%
  addTiles() %>%
  setView(map_centre[1], map_centre[2], 13) %>%
  addRasterImage(map.agb, colors = pal.biomass) %>%
  addLegend(pal = pal.biomass, values = c(0, 70), opacity = 1)
m.agb
```

Total BGB in kg/ha: `r round(biomass.bgb$sum, 2) %>% format(scientific = FALSE)`
BGB map derived from general relationship between AGB and BGB.

```{r mapBGB}
#| echo: false
#| error: false
#| message: false
#| warning: false

m.bgb <- leaflet() %>%
  addTiles() %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  addRasterImage(map.bgb, colors = pal.biomass) %>%
  addLegend(pal = pal.biomass, values = c(0, 70), opacity = 1)
m.bgb
```


# Carbon stock prediction

Maps and estimated total stock in Ria de Aveiro. 

```{r predictedcarbonstock}
#| echo: false
#| error: false
#| message: false
#| warning: false

## g/m2 * g/g --> g as output
aveiro.carbon.leaf.factor <- 0.2870144
derive.carbon.leaf <- function(map.biomass.leaf) {
  map.biomass.leaf * aveiro.carbon.leaf.factor
}

map.agc <- derive.carbon.leaf(map.agb)

pal.carbon <- colorNumeric(c('white','black'), domain = c(0, 50), na.color = 'transparent')

## g/m2 * g/g --> g as output
aveiro.carbon.root.factor <- 0.252088
derive.carbon.root <- function(map.biomass.root) {
  map.biomass.root * aveiro.carbon.root.factor
}

map.bgc <- derive.carbon.leaf(map.bgb)

## conversion to kg (divide by 1000)
carbon.agc <- global(map.agc, fun = 'sum', na.rm = TRUE) / 1000
carbon.bgc <- global(map.bgc, fun = 'sum', na.rm = TRUE) / 1000
```


Average leaf carbon content factor for Aveiro = `r round(aveiro.carbon.leaf.factor, 3)`

AGB carbon stock in Aveiro (kg) = `r round(carbon.agc, 3)`

```{r mapAGC}
#| echo: false
#| error: false
#| message: false
#| warning: false

m.agc <- leaflet() %>%
  addTiles() %>%
  setView(map_centre[1], map_centre[2], 13) %>%
  addRasterImage(map.agc, colors = pal.carbon) %>%
  addLegend(pal = pal.biomass, values = c(0, 50), opacity = 1)
m.agc
```

Average root carbon content factor for Aveiro = `r round(aveiro.carbon.root.factor, 3)`

BGB carbon stock in Aveiro (kg) = `r round(carbon.bgc, 3)`

```{r mapBGC}
#| echo: false
#| error: false
#| message: false
#| warning: false

m.bgc <- leaflet() %>%
  addTiles() %>%
  setView(map_centre[1], map_centre[2], 13) %>%
  addRasterImage(map.bgc, colors = pal.carbon) %>%
  addLegend(pal = pal.biomass, values = c(0, 50), opacity = 1)
m.bgc
```

<!-- Soil carbon is still a missing gap -->


# (Optional) Phenology

Seagrass phenology, biomass, carbon stock






