---
title: "Predicted seagrass carbon map at Ria de Aveiro for 2024 and 2025"
author: "Lee, Chengfa Benjamin"
date: "`r Sys.Date()`"
output: bookdown::html_document2
bibliography: references.bib 
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(magrittr)
library(sf)
library(dplyr)
library(leaflet)
library(terra)
# library(readxl)
# library(tidyr)
# library(ggplot2)
# library(ggiraph)
# library(ggpubr)
# library(patchwork)
```

This page documents the results of the _Zostera noltii_ seagrass carbon mapping work at Ria de Aviero. This includes work done in:

* [The fieldtrip of Aveiro 2025](https://sigoiry.github.io/Spectra_Aveiro_2025/) by [Simon Oiry](https://github.com/SigOiry)
* [Coring Carbon Seagrass](https://sigoiry.github.io/Coring_Carbon_Seagrasses/) by Simon Oiry
* [Exploring the relationship between seagrass reflectance and biomass](https://leechengfa.github.io/Seagrass_reflectance_biomass_2025.github.io/)

All analyses were performed in R version 4.5.2 [@Rversion].
All work belong to the [REWRITE project](https://rewriteproject.eu/).

# Study Location

This page focuses on Ria de Aveiro, a [demonstrator site for the REWRITE PROJECT](https://rewriteproject.eu/demonstrator-network/ria-de-aveiro). A natural site and rewilded site was designated in the lagoon and seagrass meadows were found in both sites. Its seagrass meadow was estimated to be 3.52 km^2^ during the peak season of 2024 [@davies2024].

The sites were sampled according to a grid, and additional stations were sampled randomly within the grid.

```{r samplingDesign}
#| echo: false
#| error: false
#| message: false
#| warning: false

stations <- read_sf('data/shp/Points.shp') %>%
  filter(!is.na(seagrassA)) %>%
  select(name, type, site, habitat, longitude, latitude)

map_centre <- stations %>% 
  st_union() %>%
  st_centroid() %>%
  st_coordinates()

pal.shp <- colorFactor(c('darkgreen','pink4'), domain = c('natural','rewilded'))

m.shp <- leaflet(stations) %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(map_centre[1], map_centre[2], 14) %>%
  addCircleMarkers(radius = 2, color = ~pal.shp(site), fillOpacity = 1, label = ~name) %>%
  addLegend(pal = pal.shp, values = ~site, opacity = 1)

m.shp
```

# ICE CREAMS Model prediction

ICE CREAMS classification

```{r icecreams}
#| echo: false
#| error: false
#| message: false
#| warning: false

import.icecreams <- function(image.path) {
  map.ice <- rast(image.path)
    map.ice[map.ice == 0 | map.ice == -1] <- NA
    
  return(map.ice)
}

pal.icecreams <- c('#ded4c3','#926e62','#c6c97c','#8ca271','#ddac74','#a9665e','#ca474f','#9ec8c7','#b8dc70')
labels.icecreams <- c("Bare Sand","Bare Sediment","Chlorophyta","Magnoliopsida","Microphytobenthos","Phaeophyceae","Rhodophyta","Water","Xanthophyceae")

map.class <- import.icecreams('data/ice_creams_prediction/S2C_MSIL2A_20250715T113341_N0511_R080_T29TNF_20250715T165315_Pred_BFRD_V1_3.tif')
  levels(map.class$Out_Class) <- data.frame(id=1:9, cover = labels.icecreams)

pal.class <- colorFactor(pal.icecreams, domain = 1:9, na.color = "transparent")

m.class <- leaflet() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  addRasterImage(map.class, colors = pal.class) %>%
  addLegend(colors = pal.icecreams, values = 1:9, labels = labels.icecreams, opacity = 1)
m.class
```

```{r icecreamsseagrass}
#| echo: false
#| error: false
#| message: false
#| warning: false

map.class.seagrass <- map.class
  map.class.seagrass$Out_Class[map.class.seagrass$Out_Class != 4] <- NA
  levels(map.class.seagrass$Out_Class) <- data.frame(id=4, cover = labels.icecreams[4])

m.class.sg <- leaflet() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  addRasterImage(map.class.seagrass, colors = pal.icecreams[4]) %>%
  addLegend(colors = pal.icecreams[4], values = 4, labels = labels.icecreams[4], opacity = 1)
m.class.sg
```

Sentinel-2 NDVI Map

```{r ndvi}
#| echo: false
#| error: false
#| message: false
#| warning: false

import.ndvi <- function(image.path) {
  map.index <- rast(image.path)
  
  return(map.index)
}

map.ndvi <- import.ndvi('data/ndvi/S2C_MSIL2A_20250715T113341_N0511_R080_T29TNF_20250715T165315_ndvi.tif')

pal.ndvi <- colorNumeric(c('white','green'), domain = c(0, 1), na.color = "transparent")

## function copied from stackoverflow
## https://stackoverflow.com/questions/40276569/reverse-order-in-r-leaflet-continuous-legend
addLegend_decreasing <- function (map, position = c("topright", "bottomright", "bottomleft","topleft"),
                                  pal, values, na.label = "NA", bins = 7, colors, 
                                  opacity = 0.5, labels = NULL, labFormat = labelFormat(), 
                                  title = NULL, className = "info legend", layerId = NULL, 
                                  group = NULL, data = getMapData(map), decreasing = FALSE) {
  
        position <- match.arg(position)
        type <- "unknown"
        na.color <- NULL
        extra <- NULL
        if (!missing(pal)) {
            if (!missing(colors)) 
                stop("You must provide either 'pal' or 'colors' (not both)")
            if (missing(title) && inherits(values, "formula")) 
                title <- deparse(values[[2]])
            values <- evalFormula(values, data)
            type <- attr(pal, "colorType", exact = TRUE)
            args <- attr(pal, "colorArgs", exact = TRUE)
            na.color <- args$na.color
            if (!is.null(na.color) && col2rgb(na.color, alpha = TRUE)[[4]] == 
                    0) {
                na.color <- NULL
            }
            if (type != "numeric" && !missing(bins)) 
                warning("'bins' is ignored because the palette type is not numeric")
            if (type == "numeric") {
                cuts <- if (length(bins) == 1) 
                    pretty(values, bins)
                else bins   
                if (length(bins) > 2) 
                    if (!all(abs(diff(bins, differences = 2)) <= 
                                     sqrt(.Machine$double.eps))) 
                        stop("The vector of breaks 'bins' must be equally spaced")
                n <- length(cuts)
                r <- range(values, na.rm = TRUE)
                cuts <- cuts[cuts >= r[1] & cuts <= r[2]]
                n <- length(cuts)
                p <- (cuts - r[1])/(r[2] - r[1])
                extra <- list(p_1 = p[1], p_n = p[n])
                p <- c("", paste0(100 * p, "%"), "")
                if (decreasing == TRUE){
                    colors <- pal(rev(c(r[1], cuts, r[2])))
                    labels <- rev(labFormat(type = "numeric", cuts))
                }else{
                    colors <- pal(c(r[1], cuts, r[2]))
                    labels <- rev(labFormat(type = "numeric", cuts))
                }
                colors <- paste(colors, p, sep = " ", collapse = ", ")
            }
            else if (type == "bin") {
                cuts <- args$bins
                n <- length(cuts)
                mids <- (cuts[-1] + cuts[-n])/2
                if (decreasing == TRUE){
                    colors <- pal(rev(mids))
                    labels <- rev(labFormat(type = "bin", cuts))
                }else{
                    colors <- pal(mids)
                    labels <- labFormat(type = "bin", cuts)
                }
            }
            else if (type == "quantile") {
                p <- args$probs
                n <- length(p)
                cuts <- quantile(values, probs = p, na.rm = TRUE)
                mids <- quantile(values, probs = (p[-1] + p[-n])/2, na.rm = TRUE)
                if (decreasing == TRUE){
                    colors <- pal(rev(mids))
                    labels <- rev(labFormat(type = "quantile", cuts, p))
                }else{
                    colors <- pal(mids)
                    labels <- labFormat(type = "quantile", cuts, p)
                }
            }
            else if (type == "factor") {
                v <- sort(unique(na.omit(values)))
                colors <- pal(v)
                labels <- labFormat(type = "factor", v)
                if (decreasing == TRUE){
                    colors <- pal(rev(v))
                    labels <- rev(labFormat(type = "factor", v))
                }else{
                    colors <- pal(v)
                    labels <- labFormat(type = "factor", v)
                }
            }
            else stop("Palette function not supported")
            if (!any(is.na(values))) 
                na.color <- NULL
        }
        else {
            if (length(colors) != length(labels)) 
                stop("'colors' and 'labels' must be of the same length")
        }
        legend <- list(colors = I(unname(colors)), labels = I(unname(labels)), 
                                     na_color = na.color, na_label = na.label, opacity = opacity, 
                                     position = position, type = type, title = title, extra = extra, 
                                     layerId = layerId, className = className, group = group)
        invokeMethod(map, data, "addLegend", legend)
        }

m.ndvi <- leaflet() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  addRasterImage(map.ndvi, colors = pal.ndvi) %>%
  addLegend_decreasing(pal = pal.ndvi, values = c(0, 1), opacity = 1, decreasing = TRUE)
m.ndvi
```

# Biomass prediction 

Rewilded vs Natural, Averaged results for the mapping

```{r predictedbiomass}
#| echo: false
#| error: false
#| message: false
#| warning: false

derive.agb <- function(ndvi) {
  agb <- ndvi * 140 - 9.5
    agb[agb < 0] <- 0
  
  return(mask(agb, map.class$Out_Class, inverse = TRUE, maskvalues = 4))
}

map.agb <- derive.agb(map.ndvi)

derive.bgb <- function(agb) {
  bgb <- agb * 0.69 + 9.1
    bgb[agb == 0] <- 0
  
  return(bgb)
}

map.bgb <- derive.bgb(map.agb)

## calculate AGB and BGB over unit area (g/m2) 
## conversion first to stock by multiplying by pixel area of Sentinel-2
## then conversion to kg per ha (multiply by 10 and divide by 1000)
biomass.agb <- global(map.agb, fun = 'sum', na.rm = TRUE) * 100 * 10 / 1000
biomass.bgb <- global(map.bgb, fun = 'sum', na.rm = TRUE) * 100 * 10 / 1000

## shared palette
pal.biomass <- colorNumeric(c('white','black'), domain = c(0, 70), na.color = 'transparent')
```

```{r mapAGB}
#| echo: false
#| error: false
#| message: false
#| warning: false

m.agb <- leaflet() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(map_centre[1], map_centre[2], 13) %>%
  addRasterImage(map.agb, colors = pal.biomass) %>%
  addLegend_decreasing(pal = pal.biomass, values = c(0, 70), opacity = 1, decreasing = TRUE)
m.agb
```

BGB map derived from general relationship between AGB and BGB.

```{r mapBGB}
#| echo: false
#| error: false
#| message: false
#| warning: false

m.bgb <- leaflet() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  setView(map_centre[1], map_centre[2], 12) %>%
  addRasterImage(map.bgb, colors = pal.biomass) %>%
  addLegend_decreasing(pal = pal.biomass, values = c(0, 70), opacity = 1, decreasing = TRUE)
m.bgb
```


# Carbon stock prediction

```{r predictedcarbonstock}
#| echo: false
#| error: false
#| message: false
#| warning: false

## g/m2 * g/g --> g as output
aveiro.carbon.leaf.factor <- 0.2870144
derive.carbon.leaf <- function(map.biomass.leaf) {
  map.biomass.leaf * aveiro.carbon.leaf.factor
}

map.agc <- derive.carbon.leaf(map.agb)

pal.carbon <- colorNumeric(c('white','black'), domain = c(0, 50), na.color = 'transparent')

## g/m2 * g/g --> g as output
aveiro.carbon.root.factor <- 0.252088
derive.carbon.root <- function(map.biomass.root) {
  map.biomass.root * aveiro.carbon.root.factor
}

map.bgc <- derive.carbon.leaf(map.bgb)

## conversion to ton (divide by 1,000,000)
carbon.agc <- global(map.agc, fun = 'sum', na.rm = TRUE) / 1000000
carbon.bgc <- global(map.bgc, fun = 'sum', na.rm = TRUE) / 1000000
```

<!-- Soil carbon is still a missing gap -->

```{r biomassCarbonTable}
#| echo: false
#| error: false
#| message: false
#| warning: false

table.bc <- data.frame(
    list(
      "component" = c('Aboveground', 'Belowground', 'Soil'),
      "surface_area" = rep(global(map.class.seagrass$Out_Class, fun = 'sum', na.rm = TRUE) / 4 * 100, 3) %>% unlist(),
      "biomass_per_area" = c(round(biomass.agb$sum, 2) %>% format(scientific = FALSE), round(biomass.bgb$sum, 2) %>% format(scientific = FALSE), 'NA'),
      "biomass" = c('Highly variable seagrass presence/cover', 'Consistently dense', 'NA'),
      "carbon_content" = c(round(aveiro.carbon.leaf.factor, 3), round(aveiro.carbon.root.factor, 3), 'NA'),
      "carbon_per_area" = c('agb_total_area', 'bgb_total_area', 'NA'),
      "carbon_stock" = c(round(carbon.agc$sum, 3), round(carbon.bgc$sum, 3), 'NA')
    )) 

names(table.bc) <- c("Component", "Total surface area<br>(m^2^)", "Total biomass per area<br>(kg DW/ha)", "Total biomass<br>(kg DW)", "Carbon content<br>(g C/g DW)", "Carbon stock per area<br>(kg C/m^2^)", "Total carbon stock<br>(ton)")

knitr::kable(table.bc, 
             format = "pipe",
             align = 'c',
             caption = 'Seagrass biomass and carbon at Ria de Aveiro. "DW" denotes the dry weight of the seagrass samples while "C" denotes the carbon in the seagrass samples.')
```



# Phenology

Seagrass phenology, biomass, carbon stock

```{r phenologyCalculations}
#| echo: false
#| error: false
#| message: false
#| warning: false

# phenology <- data.frame(list(
#   "tile.name" = list.files('data/ice_creams_prediction') %>%
#     gsub('_Pred_BFRD_V1_3.tif', '', .)
# ))
#   phenology$date <- lapply(phenology$tile.name, FUN = function(string) {
#     x <- string %>%
#       strsplit("_") %>%
#       unlist() %>%
#       "["(3) %>%
#       strsplit("T") %>%
#       unlist() %>%
#       "["(1)
#     
#     return(x)
#   }) %>% unlist() %>% as.Date('%Y%m%d')
#   phenology$surface_area <- NA
#   phenology$agb_per_area <- NA
#   phenology$bgb_per_area <- NA
#   phenology$agb_total <- NA
#   phenology$bgb_total <- NA
#   phenology$agc_per_area <- NA
#   phenology$bgc_per_area <- NA
#   phenology$agc_total <- NA
#   phenology$bgc_total <- NA
# 
# for (n in phenology$tile.name) {
#   pheno.class <- list.files('data/ice_creams_prediction', full.names = TRUE) %>%
#     grep(n, ., value = TRUE) %>%
#     import.icecreams() 
#     levels(pheno.class$Out_Class) <- data.frame(id=1:9, cover = labels.icecreams)
#     
#   pheno.class.seagrass <- pheno.class$Out_Class
#     pheno.class.seagrass[pheno.class.seagrass != 4] <- NA
#       levels(pheno.class.seagrass) <- data.frame(id=4, cover = labels.icecreams[4])
#   phenology$surface_area[phenology$tile.name == n] <- global(pheno.class.seagrass, fun = 'sum', na.rm = TRUE) / 4 * 100
#   
#   pheno.ndvi <- list.files('data/ndvi', full.names = TRUE) %>%
#     grep(n, ., value = TRUE) %>%
#     import.icecreams()
#   
#   
# }

```




# Reference